Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin

### 2017-10-13 paper/manuscript.tex
a bit more progress on third section of Results

\subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
We performed mutational antigenic profiling using the three broad antibodies (FI6v3, C179, and S139/1) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization. 
All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}.
We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three strain-specific antibodies (H17-L19, H17-L10, and H17-L7) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived each selection.
The correlations among replicates are in Figure~\ref{suppfig:corr}.
For the remainder of this paper, we will refer to the median antigenic effect of each mutation taken across all replicates with that antibody at that concentration.

@@

It is immediately obvious that both the three strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select antigenic mutations with large effects at multiple sites.
Specifically, Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are sites where a substantial fraction of virions with any mutation survive antibody concentrations that are far in excess of those needed to nearly complete neutralize the wildtype virus.
% Worth mentioning how some escape mutants from H17L19 have reduced fraction escape at the higher concentration, while this is not seen for S139/1?

>>>>

\subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization. 
All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}.
We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
The correlations among replicates are in Figure~\ref{suppfig:corr}.
For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.

@@

It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations.
Specifically, there are mutations that enable over a third of the virions to evade antibody at concentrations where virtually all the wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
Therefore, the virus can escape these four antibodies with the type of large-effect single amino-acid mutations that are thought to underly to influenza antigenic drift.
